Make any stupid plot you want

### 2021-01-18 draft/sections/implementation.tex
started fleshing out new figures

\documentclass[../main.tex]{subfiles}
\begin{document}

@@

Because we do not expect the user to specify every visual parameter, we define a class of required visual variables that are intrinsic to S. %%need to revise for formality but
What I mean here is that you cannot construct the mark w/o this - position for scatter \& line, values for heatmap.

@@

which we implement building on the \texttt{matplotlib.collections} API:

\begin{minted}{python}
    class Point(mcollections.Collection):
        # this is the visual fiber 
        required = {'x', 'y'}
        optional = {'facecolors', 's'} 
        def __init__(self, data, transforms, *args, **kwargs):
            # check that the constraints on Q can be met  
        def draw(self, renderer, *args, **kwargs):
            # assemble the glyph
            super().draw(renderer, *args, **kwargs)
    \end{minted}

(subject to change). Unlike equation~\ref{eq:scatter}, we pass also pass a reference to the data so that getting the data can be fully curried until absolutely necessary. The \texttt{*args} and \texttt{**kwargs} are artifacts of using Matplotlib as a base. We pass in the $\mu$ as the dictionary "transforms" mostly for readability, which is also why we define the visual fiber as class attributes.

Inside the \texttt{__init__} constructor, is a check if there is a valid mapping between $S$ and $K$.  Line 7-8 check if the user has passed in the required $\mu$ and whether the $\nu$ functions are valid for the input data.

\begin{minted}[lineno]{python}
def __init__(self, data, transforms, *args, **kwargs):
    super().__init__(*args, **kwargs)
    # check that the data you're trying to transform 
    # has a way to provide vertex data  
    assert 'vertex' in data.FB.K['tables']
    # check that you've given the required parameters
    utils.check_constraints(Point, transforms.keys())
    utils.validate_transforms(data.FB.F, transforms)

@@

Lines 9-10 attach the data and transforms to the $Point$ object as a concession to the Matplotlib architecture that seperates drawing from creation. In \texttt{draw}, the attributes of the graphic are set to the different sections of the visual fiber bundle.

\begin{minted}[lineno]{python}
def draw(self, renderer, *args, **kwargs):
    view = self.data.view('vertex') #resolve to size

@@

    self.set_facecolors(visual['facecolors'])
    super().draw(renderer, *args, **kwargs)
\end{minted}
Line 1 pulls back into the data bundle and returns $\tau = \{\tau_0, \ldots, tau_{n}\}$. Line 3 applies the transforms $\nu$ to $\tau$ to build the visual bundle $V$.  Line 8 sets the abstract segment marks with the visual characteristics. Line 10 passes this information to the renderer, acting like a $rho$.

The visual fiber bundle is specified as a dictionary $\{\mbox{parameter}: (\mbox{variable}, \nu), \mbox{parameter}:\mbox{value}\}$ 
\begin{minted}[lineno]{python}

@@

where the $\nu$ functions are represented as classes so that parameters can be curried. For example, \texttt{color.Categorical} is a whole set of $\nu$ functions explicitly tuned to specific categorical mappings. A pure functional approach would mean the user would have to write new functions

\begin{minted}[lineno]{python}
def color_true_deepink_false_deepink(value):
    return {'true':'deeppink', 'false':'deepskyblue'}[value]

@@

\begin{minted}[lineno]{python}
     'facecolors': ('v3', (color.categorical, {'true':'deeppink', 'false':'deepskyblue'}))
\end{minted}

but that doesn't allow for as easy reuse? (I should probably actually just switch to this form) The \textt{color.categorical} function is implemented as 
\begin{minted}[lineno]{python}
class Categorical:
    def __init__(self, mapping):
        """goal of init is to store parameters that would otherwise be
        curried higher level function"""
        assert(mcolors.is_color_like(color) for color in mapping.values())
        self._mapping = mapping

@@

where the \texttt{convert} function converts the input value to the internal normalized form Matplotlib expects. The `\texttt{validate}' function checks which monoid actions the transform supports; optimally this method should be replaced by functions that check the properties, for example a \texttt{is_ordinal} method that checks that monotonicity is preserved.

#does the data even matter?

>>>>

\documentclass[../main.tex]{subfiles}

\begin{document}

@@

Because it is impractical for the user to specify every possible visual parameter, we define a required set of visual variables that must be explicitly specified to generate distinguishable marks.

@@

which we implement building on the \mintinline{python}{matplotlib.collections} API:
%%rewrite that Point is a fully parameterized A  with the transforms dict - callable objects are a way to do tractable currying, can use __call__ same as convert - don't mutate internal state of object

\begin{minted}[linenos]{python}
    class Point(mcollections.Collection):
        # this is the visual fiber 
        required = {'x', 'y'}
        optional = {'facecolors', 's'} 
        def __init__(self, data, transforms, *args, **kwargs):
            # check that the constraints on Q can be met  
        def draw(self, renderer, *args, **kwargs):
            # assemble the glyph
            super().draw(renderer, *args, **kwargs)
    \end{minted}

(subject to change). Unlike equation~\ref{eq:scatter}, we pass also pass a reference to the data so that getting the data can be fully curried until absolutely necessary. The \mintinline{python}{*args} and \mintinline{python}{**kwargs} are artifacts of using Matplotlib as a base. We pass in the $\mu$ as the dictionary "transforms" mostly for readability, which is also why we define the visual fiber as class attributes.

Inside the \mintinline{python}{__init__} constructor, is a check if there is a valid mapping between $S$ and $K$.  Line 7-8 check if the user has passed in the required $\mu$ and whether the $\nu$ functions are valid for the input data.

%% semi functional psuedocode - strip down to don't care about broadcasting & simplest form of validation -
- assume that the $\nu$ are inherently valid - no error checking
- don't try to exactly encode monoid - is it a key feature you want to encode
- What is the important bits?

\begin{minted}[linenos]{python}
def __init__(self, data, transforms, *args, **kwargs):
    super().__init__(*args, **kwargs)
    # check that the data you're trying to transform 
    # has a way to provide vertex data  
    assert 'vertex' in data.FB.K['tables'] #there is no check here

    # check that you've given the required parameters
    utils.check_constraints(Point, transforms.keys())
    utils.validate_transforms(data.FB.F, transforms)

@@

Lines 9-10 attach the data and transforms to the $Point$ object as a concession to the Matplotlib architecture that seperates drawing from creation. In \mintinline{python}{draw}, the attributes of the graphic are set to the different sections of the visual fiber bundle.

\begin{minted}[linenos]{python}
def draw(self, renderer, *args, **kwargs):
    # use \xi^{-1} because we want to pull forward the data
    # propogate information of the downselection to do that 
    # restrict sheaf to subset of data over current view - feedback at draw restriction
    view = self.data.view('vertex') #resolve to size \xi^{-1}

@@

    self.set_facecolors(visual['facecolors'])
    super().draw(renderer, *args, **kwargs)
\end{minted}

Line 1 pulls back into the data bundle and returns $\tau = \{\tau_0, \ldots, tau_{n}\}$. Line 3 applies the transforms $\nu$ to $\tau$ to build the visual bundle $V$.  Line 8 sets the abstract segment marks with the visual characteristics. Line 10 passes this information to the renderer, acting like a $rho$.

The visual fiber bundle is specified as a dictionary $\{\mbox{parameter}: (\mbox{variable}, \nu), \mbox{parameter}:\mbox{value}\}$ 
\begin{minted}[linenos]{python}

@@

where the $\nu$ functions are represented as classes so that parameters can be curried. For example, \mintinline{python}{color.Categorical} is a whole set of $\nu$ functions explicitly tuned to specific categorical mappings. A pure functional approach would mean the user would have to write new functions

\begin{minted}[linenos]{python}
def color_true_deepink_false_deepink(value):
    return {'true':'deeppink', 'false':'deepskyblue'}[value]

@@

\begin{minted}[linenos]{python}
     'facecolors': ('v3', (color.categorical, {'true':'deeppink', 'false':'deepskyblue'}))
\end{minted}

but that doesn't allow for as easy reuse? (I should probably actually just switch to this form) The \mintinline{python}{color.categorical} function is implemented as

%%interface important, not implementation
\begin{minted}[linenos]{python}
class Categorical:
    def __init__(self, mapping):
        """goal of init is to store parameters that would otherwise be
        curried higher level function"""
        assert(mcolors.is_color_like(color) for color in mapping.values())
        self._mapping = mapping

@@

where the \mintinline{python}{convert} function converts the input value to the internal normalized form Matplotlib expects. The \mintinline{python}{validate} function checks which monoid actions the transform supports; optimally this method should be replaced by functions that check the properties, for example a \mintinline{python}{is_ordinal} method that checks that monotonicity is preserved.

%does the data even matter?

### 2021-01-18 draft/sections/math.tex
started fleshing out new figures

\subsubsection{Sheaf and Stalk}
%%% sheaf is how we break up the base space 
Often a graphic may need to be updated with live data or support zooming in on a segement of the dataset; to support working with a subset of data, we can use the sheaf $\mathcal{O}(E)$

@@

constraint: inputs go to same output means changes to inputs mean same changes to output

@@

@@

\begin{figure}[H]
    \label{fig:simplex}
    \includegraphics{figures/math/simplex.png}
    \caption{Simplices can encode the connectivity of the data, from fully disconnected (0 simplex) observations to all observations are connected to at least 3 other observations}
\end{figure}

@@

\subparagraph{Example}
\textcolor{blue}{Need to sketch and write this example}\\
chopping up a torus maybe? talk about how that gets unpacked into triangles and then into vertices
\end{document}

>>>>

\subsubsection{Sheaf and Stalk}
\label{sec:sheaf_stalk}
%%% sheaf is how we break up the base space 
Often a graphic may need to be updated with live data or support zooming in on a segement of the dataset; to support working with a subset of data, we can use the sheaf $\mathcal{O}(E)$

@@

constraint: inputs go to same o
utput means changes to inputs mean same changes to output

@@

\subsubsection{Sample Qs}
In this section we formulate the minimal Q that will generate distinguishable graphical marks: non-overlapping scatter points, a non-infinitely thin line, and a simple heatmap.

\paragraph{Q: scatter plot}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/scatter.png}
    \label{fig:artist_scatter}
\end{figure}

\begin{equation}
Q(xpos, ypos)(\alpha_{1}, \alpha_{2}) 
\end{equation}

Given a defaut color of black, $\rho_{RGB} = (0,0,0)$. The position of this swatch of color can be computed relative to the location on the disc $S_{i}$ as shown in figure~\ref{fig:artist_scatter}:
\begin{align}
x &= size\bullet \alpha_1 \bullet \cos(\alpha_2) + xpos\\
y &= size\bullet \alpha_1 \bullet \sin(\alpha_2) + ypos
\end{align}

such that $\rho(s) = (x, y, 0, 0, 0)$ where $s$ is the region in $H$.

\paragraph{Q: line plot}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/line.png}
    \label{fig:artist_line}
\end{figure}

The line plot shown in fig~\ref{fig:artist_scatter} exemplifies the need for the jet discussed in section~\ref{sec:sheaf_stalk}

\begin{equation}
    Q(xpos, \hat{n_{1}}, ypos, \hat{n_{2}})(\alpha_1, \alpha_2)
\end{equation}

where the magnitude of the thickness is 
\begin{equation}
    \lvert n \rvert = \sqrt{{n_{1}}^2 + {n_{2}}^2}
\end{equation}
such that the components are 
\begin{equation}
    \hat{n_{1}} = \frac{n_1}{\lvert n \rvert}, \; \hat{n_{2}} = \frac{n_2}{\lvert n \rvert}
\end{equation}

which yields components of $\rho(s)$:
\begin{align}
 x = xpos(\xi(\alpha_1)) &+ \alpha_2\hat(n_1)(\xi(\alpha_1)) \\
 y = ypos(\xi(\alpha_1)) &+ \alpha_2\hat(n_2)(\xi(\alpha_2)) 
\end{align}

\paragraph{Q: heatmap}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/heatmap.png}
    \label{fig:artist_heatmap}
\end{figure}
The heatmap in figure~\ref{fig:artist_heatmap} 
\begin{equation}
Q(xpos, ypos, color)
\end{equation}
has in the simple case a direct lookup into $K$ to obtain the $\mu = (x,y,c)$ values that are mapped into 
\begin{align}
D_{RGB} &= color(\xi(\alpha_1, \alpha_2))
D_x & = xpos(\xi(\alpha_1, \alpha_2))
D_y &= ypos(\xi(\alpha_1, \alpha_2))
\end{align}
through $\rho$.

@@

\begin{figure}[h!]
    \label{fig:simplex}
    \includegraphics{figures/math/simplex.png}
    \caption{Simplices can encode the connectivity of the data, from fully disconnected (0 simplex) observations to all observations are connected to at least 3 other observations}
\end{figure}

@@

\subparagraph{Example}
\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/triangle_torus.png}
    \label{fig:triangle_torus}
\end{figure}

Given data that lies on a toroidal space as shown in figure~\ref{fig:triangle_torus}, in our model this structure can be implemented as a simplacial complex of two 2-simplexes. We unfold the torus into the two triangles that compose the square in figure~\ref{fig:artist_heatmap} and encode that as K. We also constrain the functions on $A$, $A^\prime$, $B$ and $B^\prime$ such that $A$ can be glued to $A^\prime$ and $B$ to $B^\prime$ to reconstruct the torus.

\end{document}
